Adaptive Scheduling of Scientific Workflows\\ \thanks{Identify applicable funding agency here. If none, delete this.}

### 2024-05-02 main.tex
WIP. various schanges, incl results

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{algorithmicx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{xspace}

@@

\newcommand{\algo}[1]{\textsc{#1}}
\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
\newcommand{\criticalpath}{\mathcal{P}}
\newcommand{\parents}[1]{\,\Pi_{#1}}
\newcommand{\children}[1]{\,C_{#1}}
\newcommand{\cluster}{\,\mathcal{S}}
\newcommand{\daghetpart}{\algo{DagHetPart}\xspace}
\newcommand{\dagmem}{\algo{DagHetMem}\xspace}

@@

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{algorithmicx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{numprint}

@@

\newcommand{\algo}[1]{\textsc{#1}}
\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
\newcommand{\criticalpath}{\mathcal{P}}
\newcommand{\parents}[1]{\,\Pi_{#1}}
\newcommand{\children}[1]{\,C_{#1}}
\newcommand{\cluster}{\,\mathcal{S}}

\newcommand{\heftmm}{\algo{HEFTM\_MM}\xspace}
\newcommand{\heftbl}{\algo{HEFTM\_BL}\xspace}
\newcommand{\heftblc}{\algo{HEFTM\_BLC}\xspace}

@@

    \section{Experimental evaluation}
    \subsection{Setup}
    \label{sec:setup}

    All algorithms are implemented in C++ and compiled with g++ (v.11.2.0).
    The experiments are executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
    and CentOS 8 as OS.
    Code, input data, and experiment scripts are available for review under \url{TODO:insert link}.

    Next, we describe the set of workflows used in the evaluation and then the clusters on which the
    workflows are scheduled.

    \subsubsection{Workflow instances}
    The input data for the experiments consists of two sets of workflows: real-world workflows
    obtained from~\cite{ewels2020nf} and workflows obtained by simulating real-world workflows
    with the WFGen generator~\cite{COLEMAN202216}.
    First, we discuss how the graph topology is generated,
    then we focus on the weights associated to tasks and edges.

    \paragraph{Workflow graphs}
    For real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
    (and not actual tasks); that is why we removed them.

    For the simulated workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
    the desired number of tasks.
    The model workflows are described on the WFGen website, and we used the following ones:
    1000Genome, BLAST, BWA, Epigenomics, Montage, Seismology, and SoyKB.
    Other models could not be generated without errors.
%
    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
    We divide the workflows into three groups by size: small ones with up to \numprint{8000} tasks, middle
    ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.
    Note that for some workflow models, such as SoyKB or Montage, only a subset of the workflow sizes could be generated,
    because the workflow generator either took a disproportionally long time or yielded errors.

    Overall, this yields four workflow types, denoted by real, small, mid, and big.

    \paragraph{Generation of task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.
    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    To avoid underestimation, we take the maximum among all values from different runs on the same cluster node.
    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it a weight of~1.
%
    Because the historical data contains absolute measured values and the cluster node information is a relative value,
    we normalize all values extracted from the historical data by the smallest one.
    This way, task memory weights fit into the memory of the cluster nodes.
    Additionally, this way, the tasks without historical data receive less insignificant values compared to tasks with historical data.

    For the simulated workflows, we generate random execution times and memory weights for each task as well as
    edge weights for each precedence constraint.
    We generate uniformly distributed values between $1$ and $10$ for edge weights,
    $1$ and $1000$ for the workloads,
    and $1$ and $192$ for memory weights.
    When doing so, we try to mimic the weights observed in the historical data, hence \eg the low lower bounds for the
    workloads.

    \subsubsection{Target computing systems}
    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to six, thus having 36 processors in total. % the whole cluster.

        Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
        ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines; ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and large memory
        (see Table~\ref{tab:procs} in Appendix~\ref{sub:appx-cluster-config}~\cite{daghetpart_full_version}).
    Note that, by nature,
    memory sizes are normalized values, relative to each other, too.
    Therefore, we additionally normalize memory weights of real-world workflows
    to the maximum size of 192 to make sure they fit.
    For simulated workflows, we increase memory sizes proportionally until the task
    with the biggest memory requirement still
    has a processor it could be executed on.

    To test various settings, we also vary the cluster configuration and consider
    variants of the default cluster presented above:\\
    $\bullet$ {\em Small} and {\em large} clusters. While the default cluster consists of 36 nodes (6 of each kind), we
    consider a small cluster with one processor of each kind ($6$ processors in total), and a
    large cluster with ten processors of each kind ($60$ processors in total). \skug{leaving the big cluster inside for
    the case we do want to include it}\\
    $\bullet$ {\em Fat-and-thin} cluster. In this cluster, the fastest processors have the smallest amount of memory (thin),
    while the slowest processors have the largest memories (fat).
    The overall amount of memory in the cluster, as well as the overall speed of all processors remain the same as in the
            {\em normal} cluster.
%\end{itemize}

    \subsection{Results}
    \subsubsection{Normal cluster}

    On the normal cluster, the baseline produces invalid results in $40$\% of small workflows and $66.6$\% of middle-sized
    and big workflows.
    In most cases, the failures of the baseline can be traced to choosing the processor with not enough memory when several
    options are available.
    Due to not taking memories into account, the baseline chooses a processor with already full memory, while there can
    be another processor with the same speed that still has enough memory.
    In this case, \heftbl produces the same makespan while producing a valid solution.
    The heuristics produce valid results by default and also manage to find a solution in all cases (on all workflow sizes).

    On the normal cluster and among small workflows, the makespans produced by \heftbl and \heftblc are $1$\% worse than the baseline,
    \heftmm $19$\% worse.
    For middle-sized ones \heftmm is $12.3$\% worse, the rest is also $1-2$\% worse.
    For the big workflows however, the results of \heftbl and \heftblc are $<1$\% worse and \heftmm is only $9$\% worse.
    The heuristics pay off more on larger workflows.
    \subsubsection{Memory usage}
    \subsubsection{Impact of cluster size}
    On a cluster with only 6 processors (1 of each kind), the baseline makes more failures: from $43.4$\%  (small workflows)
    to $88.9$\% (mid-sized) and $94.4$\% (big workflows).
    These values also become more unattainable: the makespans heuristics produce are worse by $3-12$\% (\heftbl) to $12-52$\%
    (\heftmm).
    However, these comparisons are less expressive, because most of baseline solutions are invalid.
    \skug{TODO: add numbers about failures on the small cluster with the fixed bug}

    \subsubsection{Impact of cluster composition}
    On the {\em fat-and-thin} cluster, the baseline produces $60$\% invalid results on small workflows, $83.3$\% on
    middle-sized and $88.9$\% on big workflows, and therefore performs worse than on the normal cluster.

    The relative makespans are $14-29$\% worse for small workflows, $22-40$\% worse on middle-sized workflows and $16-35$\%
    on big workflows.
    \wrt failures, all heuristics again manage to schedule all workflows.
    \subsubsection{Behavior of workflow types when scaling workflow size}
    \subsubsection{Running times}
    The geometric means of runtimes of all heuristics on small workflows lie between $0.9$ and $2.2$ seconds.
    For middle-sized workflows, the runtimes for \heftbl and \heftblc are close to $2$ minutes, while \heftmm
    takes $5.1$ minutes.
    For the big workflows, again, \heftbl and \heftblc show a better runtime ($5$ and $5.4$ minutes, respectively).
    \heftmm takes $13.9$ minutes on average.
    \subsubsection{Summary}
